---
title: "ClusteringAlgorithm1"
author: "Mandy Liesch"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
library(tree)
library(maptree)
``` 

```{r}
oneRiver<-read.csv('OneRiver1.csv')
NASS2017<-read.csv('NASS2017.csv')
countyWater<-read.csv('countyWaterViol.csv')
totalData<-merge(oneRiver, NASS2017, by.x='FIPSCODE', by.y='FIPS', all.x=TRUE)
totalData<-merge(totalData, countyWater, by.x='FIPSCODE', by.y= 'GEOINT', all.x=TRUE)

presEJ<-read.csv("presEJ.csv")
totalData<-merge(totalData, presEJ, by.x='FIPSCODE', by.y= 'GEOID', all=TRUE)

futEJ<-read.csv("futureEJ.csv")
totalData<-merge(totalData, futEJ, by.x='FIPSCODE', by.y= 'GEOID', all=TRUE)

NASSTrim<-NASS2017[-c(2:4, 24:27, 34:43, 45, 48:49, 56:59, 63:64, 66, 69:77, 143:144, 146:163)]

NASSEconTrim<-NASSTrim[-c(21:105, 108:110)]


row.names(NASSEconTrim) <- NASSEconTrim$GEOID
#remove the data not needed for clustering
clustData<-NASSEconTrim
row.names(clustData) <- clustData$FIPS
clustData<-clustData[2:20]

#change market value to numeric
clustData$MARKET_VALUE<-as.numeric(gsub(",", "", clustData$MARKET_VALUE))

clustData$AN_OPER<-as.numeric(gsub(",", "", clustData$AN_OPER))

clustData$AN_OPER<-as.numeric(gsub(",", "", clustData$AN_OPER))

clustData[1:21] <- lapply(clustData[1:21], as.numeric)
clustData[is.na(clustData)] <- 0
#Scale the data
normal<-scale(clustData)

dist_mat <- dist(normal, method = 'euclidean')

hc1 <- hclust(dist_mat, method = "ward.D" )

plot(hc1, cex = 0.25, hang = -1)

m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(dist_mat, method = x)$ac
}

map_dbl(m, ac)

hc4 <- diana(dist_mat)

hc4$dc

# Ward's method
hc5 <- hclust(dist_mat, method = "ward.D" )

# Cut tree into 4 groups
sub_grp <- cutree(hc5, k = 4)
plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 4, border = 2:5)




# Number of members in each cluster
table(sub_grp)

clustData<-clustData %>% mutate(cluster = sub_grp)

fviz_cluster(list(data = clustData, cluster = sub_grp))

fviz_nbclust(clustData, FUN = hcut, method = "wss")

```

```{r}
#Set up for the categorical tree
clustData$cluster<-as.factor(clustData$cluster)
tree.cluster = tree(cluster~., data=clustData)

summary(tree.cluster)
draw.tree(tree.cluster,cex=0.35)

tree.pred = predict(tree.cluster, clustData, type="class")

cv.cluster = cv.tree(tree.cluster, FUN = prune.misclass)
cv.cluster

prune.cluster = prune.misclass(tree.cluster, best = 12)
draw.tree(prune.cluster, cex=0.35)
text(prune.cluster, pretty=0)

#Split into test and training
set.seed(101)
train=sample(1:nrow(clustData), 800)

tree.cluster = tree(cluster~., clustData, subset=train)

plot(tree.cluster)
text(tree.cluster, pretty=0)

tree.pred = predict(tree.cluster, clustData[-train,], type="class")

with(clustData[-train,], table(tree.pred, cluster))

cv.cluster = cv.tree(tree.cluster, FUN = prune.misclass)
cv.cluster

#9 or 10 models
prune.cluster = prune.misclass(tree.cluster, best = 8)
plot(prune.cluster)
text(prune.cluster, pretty=0)
```

```{r}
clustData<-NASSEconTrim
row.names(clustData) <- clustData$FIPS
clustData<-clustData[2:20]
clusters<-clustData$cluster
clustData$clusters<-clusters

pretrimmedClust<-clustData[c(1, 20)]

pretrimmedClust$MARKET_VALUE<-as.numeric(gsub(",", "", pretrimmedClust$MARKET_VALUE))


pretrimmedClust$clusters<-as.factor(pretrimmedClust$clusters)
tree.trimcluster = tree(clusters~., data=pretrimmedClust)

summary(tree.trimcluster)
draw.tree(tree.trimcluster,cex=0.35)

```

```{r}

NASSAreaTrim<-NASSTrim[-c(3:35)]
NASSAreaTrim<-NASSAreaTrim[-c(2, 7:8, 73:77)]

row.names(NASSAreaTrim) <- NASSAreaTrim$FIPS

#remove the data not needed for clustering
clustAreaData<-NASSAreaTrim
row.names(clustAreaData) <- clustAreaData$FIPS
clustAreaData<-clustAreaData[2:69]
clustAreaData[is.na(clustAreaData)] <- 0

#Scale the data
normal<-scale(clustAreaData)

dist_mat <- dist(normal, method = 'euclidean')

m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(dist_mat, method = x)$ac
}

map_dbl(m, ac)


hc4 <- diana(dist_mat)

hc4$dc

hc5 <- hclust(dist_mat, method = "ward.D" )

# Number of members in each cluster
table(sub_grp)

clustAreaData<-clustAreaData %>% mutate(cluster = sub_grp)

fviz_cluster(list(data = clustAreaData, cluster = sub_grp))

fviz_nbclust(clustAreaData, FUN = hcut, method = "wss")


clustAreaData$cluster<-as.factor(clustAreaData$cluster)
tree.cluster = tree(cluster~., data=clustAreaData)

summary(tree.cluster)
draw.tree(tree.cluster,cex=0.35)

cv.cluster = cv.tree(tree.cluster, FUN = prune.misclass)
cv.cluster

prune.cluster = prune.misclass(tree.cluster, best = 18)
draw.tree(prune.cluster,cex=0.35)
text(prune.cluster, pretty=0)
