---
title: "ClusteringAlgorithm1"
author: "Mandy Liesch"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
``` 

```{r}
oneRiver<-read.csv('OneRiver1.csv')
NASS2017<-read.csv('NASS2017.csv')
countyWater<-read.csv('countyWaterViol.csv')
totalData<-merge(oneRiver, NASS2017, by.x='FIPSCODE', by.y='FIPS', all.x=TRUE)
totalData<-merge(totalData, countyWater, by.x='FIPSCODE', by.y= 'GEOINT', all.x=TRUE)

row.names(totalData) <- totalData$FIPSCODE
#remove the data not needed for clustering
clustData<-totalData[9:377]


#change market value to numeric
clustData$MARKET_VALUE<-as.numeric(gsub(",", "", clustData$MARKET_VALUE))

clustData[1:369] <- lapply(clustData[1:369], as.numeric)
clustData[is.na(clustData)] <- 0
#Scale the data
normal<-scale(clustData)

dist_mat <- dist(normal, method = 'euclidean')

hc1 <- hclust(dist_mat, method = "ward.D" )

plot(hc1, cex = 0.25, hang = -1)

m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(dist_mat, method = x)$ac
}

map_dbl(m, ac)

hc4 <- diana(dist_mat)

hc4$dc


hc5 <- hclust(dist_mat, method = "ward.D2" )

# Cut tree into 4 groups
sub_grp <- cutree(hc5, k = 5)
plot(hc5, cex = 0.6)
rect.hclust(hc5, k = 5, border = 2:5)

# Number of members in each cluster
table(sub_grp)

clustData<-clustData %>%
  mutate(cluster = sub_grp)

fviz_cluster(list(data = clustData, cluster = sub_grp))

fviz_nbclust(clustData, FUN = hcut, method = "wss")
