---
title: "Final Clustering Results"
author: "Mandy Liesch"
date: "10/11/2021"

---


```{r, echo=FALSE, message = FALSE, warning=FALSE}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
library(rmarkdown)
library(usethis)
library(knitr)
use_git_config(user.name="Mandy Liesch", user.email="amliesch@ncsu.edu")
```

```{r, eval=FALSE, echo=FALSE}
#code to create the rendering. this chunk is put into the code to knit and
#render the function.
rmarkdown::render("OneRiverHeirCluster.Rmd", 
                  output_format = "github_document", 
                  output_options = list(toc=TRUE, toc_depth = 3, html_preview=FALSE), 
                  output_file = "OneRiverReadMe.md")
```

# Project Objectives

There are many different ways to classify "risk" in rural areas. There are current environmental exposures, socio-economic variables, issues that result from culture shifts, population dynamic changes, changes in agricultural resilience, and future risks changes from shifting climates, crop yield changes, and environmental hazards.  

This project focused on a 12 state region: The states along the main branch of the Mississippi River basin, the Ohio, and Tennessee River.  

# Dataset Manipulation: 

## OneRiver.csv File:

There are several different datasets that were utilized to build the One River Dataset: 

### Social Determinants of Health

[Social Determinants of Health](https://www.ahrq.gov/sdoh/data-analytics/sdoh-data.html): Looking at the [documentation](https://www.ahrq.gov/sites/default/files/wysiwyg/sdohchallenge/data/sdoh_data_file_documentation.pdf), it aggregates several sources of health data. Data is availabe at both county and zip code levels. For this project, we are using the County data.  

Including:  
* American Community Survey (ACS)  
* Area Health Resources Files (AHRF)   
* Foundation for AIDS Research (amfAR)  
* U.S. Census County Business Patterns (CCBP)  
* Centers for Disease Control and Prevention Interactive Atlas of Heart
Disease and Stroke (CDC Atlas)  
* Centers for Disease Control and Prevention Wide-ranging ONline Data for
Epidemiologic Research (CDC WONDER)  
* County Health Rankings (CHR)  
* Civil Rights Data Collection (CRDC)  
* Medicare Advantage Penetration (MAP)  
* National Environmental Public Health Tracking Network (NEPHTN)  
* National Center for Health Statistics (NCHS) Urban-Rural Classification
Scheme  
* Nursing Home Compare (NHC)  
* Social Vulnerability Index (SVI)  
* U.S. Cancer Statistics (USCS)  

### Social Capital Index
Along with the social determinents of health database, the [Social Capital Index (SoCI)](https://www.researchgate.net/publication/337813421_Capturing_Bonding_Bridging_and_Linking_Social_Capital_through_Publicly_Available_Data) uses 19 indicators of public data to determine which communities are more capable of disaster resilience, based on how communities are organized and researched. They use three categories (bonding, bridging, and linking) to design an overall index by county. The base indicators that are not well represented with the above dataset were added into the Social Determinants of Health Dataset. 

### ESRI County Health Rankings
In addition, the [ESRI County Health Rankings](https://www.arcgis.com/home/item.html?id=c514eddc6d584e85bc2f90be25305fc8) that were not in the other databases were added to the csv datafile.

### Atlas of Rural America
There are still several variables that we want for analysis missing from the dataset. The USDA Economic Research Services (ERS), publishes a [Rural Atlas of Rural and Small Town America](https://www.ers.usda.gov/data-products/atlas-of-rural-and-small-town-america/). this site looks at variables of people, economic, and jobs data. The website above documents the maps and processes used to create this atlas. 

## Missing Data

### Death of Despair 
This data spanned a 5 year range from 2015-2019. Defined as Drug Overdoses, Alcohol Related Deaths, Suicides, and Homicides, getting reliable county level estimates can be difficult, especially in rural areas. The CDC suppresses any county information where there are not at least 10 of each individual type of deaths. In areas with low population density, it can take 20+ years to get to 10 of these deaths. So, to fill in these blank data rates, the [2013 Urban and Rural classification](https://www.cdc.gov/nchs/data_access/urban_rural.htm) average values were used by county type (Metropolitan Core, Micropolitan, non-core, etc.) as a substitution. This dataset deals with Non-Ignorable Missingness, meaning that the missing data trends toward rural areas. 

There are 180 counties that are suppressed with Suicide and Homicide Data:  
- 3 Large Fringe Metro: All in Indiana  
- 3 Medium Metro: All in Iowa  
- 3 Small Metro: All in Kentucky  
- 17 Micropolitan: 6 States  
- 155 Non Core Metro: 12 States   


There are 222 counties that are suppressed for Alcohol and Drug Overdoses:  
- 2 Large Fringe Metro: All in Mississippi  
- 6 Medium Metro: 2 in Arkansas, 4 in Iowa  
- 9 Small Metro: 3 States (KY, LA, MO)  
- 24 Micropolitan: 6 States  
- 183 Non Core Metro: 11 States (Not Ohio)  

### Diabetes Mellitus Deaths
Deaths from Diabetes were calculated over the five year span of 2015-2019 from the CDC Wonder Data. Like the Deaths from Despair Data, there are several counties with less than 10 deaths. 

There are 43 Counties with missing Diabetes data:   
- 2 Large Fringe Metro: Both in Indiana  
- 6 Micropolitan: Arkansas, Kentucky, and Illinois  
- 35 Non Core Metro: 7 States  

```{r, echo=FALSE}
#Load in the CDC File
CDCCorr<-read.csv("CDCCorrections.csv")
#Remove all values that don't have a metro designation.
CleanCDC<-as_tibble(CDCCorr[CDCCorr$X2013.Urbanization != "", ])
CleanCDC$State<-CleanCDC$ï..State
CleanCDC$ï..State<-NULL
CleanCDC$State.Code<-NULL
CleanCDC$Urbanization<-CleanCDC$X2013.Urbanization
CleanCDC$X2013.Urbanization<-NULL
CleanCDC$Homicide<-CleanCDC$Homicide.Rate
CleanCDC$Homicide.Rate<-NULL
#Rename the variables
table1<- CleanCDC %>%
  select(State, Urbanization, Suicide, Homicide, DrugOD, Alcohol, Diabetes)


table1<-knitr::kable(table1)

table1

```


### Firearm Fatalities
The same five year span was used to calculate all firearm fatalities from the CDC Wonder Data. 





